Data Exploratory Analyses/Plots

### 2022-12-05 DataExploratoryAnalyses.Rmd
Added specific questions to the exploratory analyses

```{r}
source('sharedVariables.R')
source('NEON_Data_DownloadAndProcess.R')
```

```{r}
library(MuMIn)
library(ggplot2)
library(mgcv)
library(ggridges)

@@

```{r}
p <- 2 #Bud burst 
  print(NEON_phenophase_names[p])
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p])
  subDat$year <- as.numeric(subDat$year)
  subDat <- left_join(subDat,allPrecipData[,-3],by=c('siteID','date','year'))
  subDat <- left_join(subDat,allTempData,by=c('siteID','date','year'))
  subDat <- left_join(subDat,allSunlightTimes,by=c('siteID','date'))
  subDat <- subDat %>% filter(year!=2022)
  subDat$Mycorrhizal.type <- as.factor(subDat$Mycorrhizal.type)

@@

    indCts <- subDatDB %>% count(individualID)
    indCts <- left_join(indCts,subDatDB[,c('siteID',"individualID")])

    ggplot(indCts,aes(x=n))+
      geom_histogram(binwidth = 1)+
      facet_wrap(~siteID,ncol=4)+
      xlab("Number of Repeated Measurements Per Individual")+
  theme_classic()

@@

```{r}
p <- 3 #Senescence
  print(NEON_phenophase_names[p])
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p])
  subDat$year <- as.numeric(subDat$year)
  subDat <- left_join(subDat,allPrecipData[,-3],by=c('siteID','date','year'))
  subDat <- left_join(subDat,allTempData,by=c('siteID','date','year'))
  subDat <- left_join(subDat,allSunlightTimes,by=c('siteID','date'))
  subDat <- subDat %>% filter(year!=2022)
  subDat$Mycorrhizal.type <- as.factor(subDat$Mycorrhizal.type)

>>>>

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE}
source('sharedVariables.R')
source('NEON_Data_DownloadAndProcess.R')
```

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE}
library(MuMIn)
library(ggplot2)
library(mgcv)
library(ggridges)

@@

#For each site-PFT combination: How many individuals of the different associations are present?

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE,fig.width=10,fig.height=15}
#Number of Individuals of AM and EcM that are present
uniqueIndividuals <- matrix(nrow=0,ncol=5)
for(p in seq_along(NEON_phenophase_names)){
  #Bud burst 
  print(NEON_phenophase_names[p])

  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat[,c('siteID','individualID','growthForm','Mycorrhizal.type','decimalLatitude')]
  subDat <- subDat %>% unique()
  uniqueIndividuals <- rbind(uniqueIndividuals,subDat)
  uniqueIndividuals <- uniqueIndividuals %>% unique()
}
uniqueIndividuals <- uniqueIndividuals[!(uniqueIndividuals$siteID%in%c("2019","2020","2021","2017","2018","2022")),]
#uniqueIndividuals <- uniqueIndividuals %>% filter(Mycorrhizal.type %in% c("AM","EcM"))
uniqueIndividuals <- uniqueIndividuals %>% filter(!is.na(growthForm))
uniqueIndividuals[uniqueIndividuals$growthForm=="Pine",'growthForm'] <- "Evergreen conifer"
uniqueIndividuals <- uniqueIndividuals %>% filter(growthForm %in% c("Deciduous broadleaf","Drought deciduous broadleaf","Evergreen broadleaf","Evergreen conifer","Forb","Graminoid","Semi-evergreen broadleaf")) #Exclude the others

uniqueIndividuals2 <- uniqueIndividuals %>% group_by(siteID,Mycorrhizal.type,growthForm,decimalLatitude) %>% summarise(n=n())

ggplot(uniqueIndividuals2,aes(fill=Mycorrhizal.type,y=n,x=fct_reorder(siteID,decimalLatitude))) + 
  geom_bar(position="dodge",stat="identity")+
  facet_wrap(~growthForm,ncol=4)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("Site ID (Ordered by Latitude)")+
  ylab("Number of Individuals")+
  coord_flip()

What I learned is that DB has the best variation between different mycorrhizal associations so that's where I should focus my comparitive studies and then do across PFTs for the overall.

#For each site-phenophase combination: how many years are individuals re-measured?
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.width=8,fig.height=10}
for(p in seq_along(NEON_phenophase_names)){
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat[,c('siteID','individualID','Mycorrhizal.type','decimalLatitude','year')]
  subDat <- subDat %>% unique()

  subDat2 <- subDat %>% group_by(siteID,Mycorrhizal.type,individualID,decimalLatitude) %>% summarise(n=n())
  if(nrow(subDat2)>2){
    print(ggplot(subDat2,aes(n,fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
            geom_density_ridges2(scale=1)+
            theme_classic()+
            theme(axis.text.x = element_text(angle=90))+
            xlab("Number Of Years Per Individual")+
            ylab("Site ID (Ordered by Latitude)")+
            ggtitle(NEON_phenophase_names[p])+
            scale_x_continuous(limits=c(0,8)))
  }
}

#For each PFT-phenophase combination: Does the Standard Deviation in DOYs and lengths between phenophases differ between EcM and AM associated plants acrross all sites and within each site?

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE,fig.width=10,fig.height=20}
for(p in seq_along(NEON_phenophase_names)){
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p])
  subDat <- subDat[,c('siteID','individualID','Mycorrhizal.type','decimalLatitude','dayOfYear','growthForm')]
  subDat <- subDat %>% filter(growthForm %in% c("Deciduous broadleaf","Drought deciduous broadleaf","Forb","Graminoid"))
  subDat2 <- subDat %>% group_by(individualID) %>% mutate(doy_sd=sd(dayOfYear,na.rm=TRUE))

  if(nrow(subDat2)>2){
    print(ggplot(subDat2,aes(doy_sd,fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
            geom_density_ridges2(scale=1)+
            theme_bw()+
            theme(axis.text.x = element_text(angle=90))+
            xlab("SD in DOY Per Individual")+
            ylab("Site ID (Ordered by Latitude)")+
            facet_wrap(~growthForm)+
            ggtitle(NEON_phenophase_names[p])+
            scale_x_continuous(limits=c(0,100)))
  }
}

#Lengths Between Phenophases:

#For each PFT-phenophase combination: Does the DOY of the timing and the lengths between phenophases differ between EcM and AM assocated plants across all sites and within each site?

```{r}

#For each individual and phenophase, what predictors seem to be the most important?

```{r}

#Do the basic GAMs work for any of the phenophase and PFTs?

```{r}

```{r,eval=FALSE}
p <- 2 #Bud burst 
  print(NEON_phenophase_names[p])
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p])
  subDat$year <- as.numeric(subDat$year)
  subDat <- left_join(subDat,allPrecipData[,-3],by=c('siteID','date','year'))
  subDat <- left_join(subDat,allTempData,by=c('siteID','date','year'))
  subDat <- left_join(subDat,allSunlightTimes,by=c('siteID','date'))
  subDat <- subDat %>% filter(year!=2022)
  subDat$Mycorrhizal.type <- as.factor(subDat$Mycorrhizal.type)

@@

@@

```{r,eval=FALSE}
p <- 3 #Senescence
  print(NEON_phenophase_names[p])
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p])
  subDat$year <- as.numeric(subDat$year)
  subDat <- left_join(subDat,allPrecipData[,-3],by=c('siteID','date','year'))
  subDat <- left_join(subDat,allTempData,by=c('siteID','date','year'))
  subDat <- left_join(subDat,allSunlightTimes,by=c('siteID','date'))
  subDat <- subDat %>% filter(year!=2022)
  subDat$Mycorrhizal.type <- as.factor(subDat$Mycorrhizal.type)
